# Results
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
read the data
```{r}
library(tidyverse)
library(gridExtra)

df<- read.csv("../winequality-red.csv")
df
```

producing boxplots for each of the variables
```{r}
draw_box <- function(df,variable)
{
  plot <- ggplot(data = df, aes(x = variable)) + 
           geom_boxplot() +
           xlab(substring(deparse(substitute(variable)),4))
            
  return(plot)
}

grid.arrange(draw_box(df, df$fixed.acidity),
             draw_box(df, df$volatile.acidity),
             draw_box(df, df$citric.acid),
             draw_box(df, df$residual.sugar),
             draw_box(df, df$chlorides),
             draw_box(df, df$free.sulfur.dioxide),
             draw_box(df, df$total.sulfur.dioxide),
             draw_box(df, df$density),
             draw_box(df, df$pH),
             draw_box(df, df$sulphates),
             draw_box(df, df$alcohol),
             draw_box(df, df$quality),
             ncol = 3)
```

find correlation for each pairs of variables.
```{r}
library(GGally)
ggpairs(df, upper = list(continuous = wrap("cor", size = 2.5)), axisLabels="none", progress = FALSE)
```


Find correlation for each pairs of variables.
```{r}
library(ggcorrplot)

cormap <- round(cor(df),1)

ggcorrplot(cormap, hc.order = TRUE, type = "lower", lab = TRUE)

```

When beginners first drink red wine, its much easier to taste the level of Sweetness and acidity of the red wine. We draw the scatterplot for residual sugar and volatile acidity with quality. We can see that only red wines with quality 4,5,6 have high and middle residual sugar and low and good quality red wines only have low residual sugar. So, if you feel the red wine is very sweet or even a little sweet, the quality red wine you drink is probably average.
We also can see that only red wine with quality 3,4,5 has high volatile acidity. So, if you feel the red wine is very sour, the quality red wine you drink is probably low or average.

```{r}
grid.arrange(
  ggplot(data = df, aes(x = quality, y = residual.sugar)) +
  geom_jitter(aes(color = residual.sugar))+
scale_color_distiller(palette = "RdYlGn")+
    ggtitle('Residual sugar with quality')+
  theme(legend.position="bottom"),
ggplot(data = df, aes(x = quality, y = volatile.acidity)) +
  geom_jitter(aes(color = volatile.acidity))+
scale_color_distiller(palette = "RdYlGn")+
    ggtitle('Volatile acidity with quality')+
  theme(legend.position="bottom"),
ncol=2
)
```

